---
title: "Project 2"
author: "Kate Sawdey, Claire Weadock, Olivia Caponecchi, Michael Zuckerman, Anna Dolce, Rubens Mondi"
date: "3/5/2023"
output:
  html_document:
    toc: true
    theme: readable
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary

[insert exec summary]

# Intro

[insert intro]

## Downloading and Prepping the Data

```{r}
#Downloading and Prepping the Data
tele <- read.csv("tele.csv", stringsAsFactors = TRUE)
summary(tele)

#We are deleting the "duration" variable because it is an after the fact measurement. We only should be using variables that we know before the call
tele$duration <- NULL

# Deleting the column X
tele$X <- NULL

# Changing pdays to a dummy and deleting pdays
tele$pdaysdummy <- ifelse(tele$pdays == 999, 0, 1)
tele$pdays <- NULL

str(tele)
```

**Additional Cleaning to think about?**

- consolidating month because has lot of levels? maybe create new variable for quarters?
- maybe consolidate education? what do basic 4/6 yrs mean?


## Getting Data Ready for Analysis

```{r}
# Using model.matrix to convert all the factors to dummy variables
# We are converting all of the factors into dummy variables as the input into knn has to be numeric

telemm <- as.data.frame(model.matrix(~.-1,tele))
str(telemm)

# Randomize the rows in the data (shuffling the rows)
set.seed(12345)
tele_random <- telemm[sample(nrow(telemm)),]

#Normalize the data
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

# we are going to normalize everything 
tele_norm <- as.data.frame(lapply(tele_random, normalize))
```


## Getting Train and Test Samples

```{r}
# Selects 10000 random rows for test data
set.seed(12345)
test_set <- sample(1:nrow(tele_norm), 10000) 
# Depending on R-version and computer, different rows may be selected. 
# If that happens, results are different. 

# Create a train set and test set
#First the predictors - all columns except the yyes column
tele_train <- tele_norm[-test_set, -match("yyes",names(tele_norm))]
tele_test <- tele_norm[test_set, -match("yyes",names(tele_norm))]

#Now the response (aka Labels) - only the yyes column
tele_train_labels <- tele_norm[-test_set, "yyes"]
tele_test_labels <- tele_norm[test_set, "yyes"]

```

> Now you are ready to build your ANN model. Feel free to modify the data load, cleaning and preparation code above as per your preference.


# Logistic Regression

as;lkfj;alsdjf;lajg;ljsag;ljsdlfj


# KNN Model


# ANN Model


# Combining Models Together


# Results and Analysis


# Conclusion